    
# Lab | Data Wrangling with R

## Objective

Practice data wrangling, cleaning, and manipulation techniques in R using the Sample Super Store dataset. By the end of this activity, you should be able to:

- Load and explore the dataset.
- Perform basic and advanced data manipulations.
- Handle missing data and duplicates.
- Aggregate and summarize data.
- Dataset Overview.

The Sample Super Store dataset contains sales, profit, and customer information for a retail store. It includes columns such as:

- Order ID, Order Date, Ship Date
- Customer ID, Customer Name, Segment
- City, State, Region, Category, Sub-Category
- Sales, Quantity, Discount, Profit

## Instructions

### Step 1: Load and Explore the Dataset

1. Load the dataset into R.
```{r}
super_store <- read.csv("dataset/Sample - Superstore.csv")
```

2. Explore the dataset using str(), head(), and summary().
```{r}
head(super_store)
```
```{r}
summary(super_store)
```


```{r}
str(super_store)

```

3. Identify the number of rows and columns.
```{r}
nrow(super_store)
ncol(super_store)
```
```{r}
colSums(is.na(super_store))
```


### Step 2: Basic Data Manipulation

1. Select the following columns: Order ID, Order Date, Customer Name, Sales, Profit.
```{r}
library(dplyr)
super_store1 <- super_store %>% select("Order.ID", "Order.Date", "Customer.Name", "Sales", "Profit")
```

2. Filter the dataset to show only orders with a profit greater than $100.
```{r}
filtered_profit <- super_store %>%
  filter(Profit > 100)
```

3. Sort the dataset by Sales in descending order.
```{r}
sorted_by_sales <- super_store %>%
  arrange(desc(Sales))
```


### Step 3: Handle Missing Data

1. Check for missing values in the dataset.
```{r}
colSums(is.na(super_store))
```

2. Replace missing values in the Postal Code column with the mode (most frequent value).
```{r}
#Replace missing values with Mode
library(DescTools)

super_store <- super_store %>%
  mutate(Postal.Code = ifelse(is.na(Postal.Code), Mode(Age, na.rm = TRUE), Postal.Code))
```

3. Remove rows with missing values in the Customer Name column.
```{r}
super_store <- super_store %>%
  filter(!is.na(Customer.Name))
```

### Step 4: Create and Modify Columns

1. Create a new column Profit_Margin as the ratio of Profit to Sales.
```{r}
super_store <- super_store %>%
  mutate(Profit_Margin= Profit/Sales)

head(super_store)
```

2. Create a new column Order_Year by extracting the year from Order Date.
3. Convert the Order.Date column to a Date data type.
```{r}
library(lubridate)

super_store <- super_store %>%
  mutate(Order.Date = mdy(Order.Date),  # Convert to Date type 
         Order_Year = year(Order.Date))  # Extract the year from Order.Date

```


```{r}
head(super_store)
```

### Step 5: Aggregating and Summarizing Data

1. Calculate the total sales and profit by Category.
```{r}
category_summary <- super_store %>%
  group_by(Category) %>%
  summarise(
    Total_Sales=sum(Sales),
    Total_Profit=sum(Profit)
  )
print(category_summary)
```

2. Find the average profit margin by Region.
```{r}
average_profit_margin <- super_store %>%
  group_by(Region) %>%
  summarise(
    Average_Profit_Margin= Mean(Profit_Margin)
    )

print(average_profit_margin)
  
```

3. Count the number of orders by Customer Segment.
```{r}
number_orders_by_segment <- super_store %>%
  group_by(Segment) %>%
  summarise(Order_count= n())

print(number_orders_by_segment)
```


## Optional

### Extra 1: Advanced Challenges

1. Identify and remove duplicate rows based on Order ID.
```{r}
#Checking for duplicate rows
duplicated_rows <- super_store[duplicated(super_store$Order.ID), ]

nrow(duplicated_rows)
```

2. Create a new column Discount_Level that categorizes discounts as "Low" (<10%), "Medium" (10-20%), and "High" (>20%).
```{r}
#Removing duplicate rows
super_store_unique <- super_store %>%
  distinct(Order.ID, .Keep_all= TRUE)
nrow(super_store_unique)
```

3. Merge the dataset with a new dataset containing regional population data (create a dummy dataset for this purpose).
```{r}
# Check unique regions in super_store
unique(super_store$Region)


```

```{r}
# Create a dummy dataset to join
population_data <- data.frame(
  Region = c("South","West","Central","East"),
  Population = c(1000000, 1500000, 2000000, 2500000)
)

print(population_data)
```

```{r}
# Join with the Super_Store dataset
super_store <- super_store %>%
  left_join(population_data, by = "Region")

# View the updated dataset
head(super_store)
```


### Extra 2: Data Visualization

Next class we will talk about data visualization, but let's see if you can pull it off on your own.

1. Create a bar plot of total sales by Category.
```{r}
#Total sales by category
sales_by_category <- super_store %>%
  group_by(Category) %>%
  summarise(
    Total_Sales= sum(Sales)
  )
head(sales_by_category)
```
```{r}
#Bar plot
library(ggplot2)
library(scales)

ggplot(sales_by_category, aes(x = Category, y = Total_Sales, fill = Category)) +
  geom_bar(stat = "identity") +
  labs(title = "Total Sales by Category",
       x = "Category",
       y = "Total Sales") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5)) +  
  scale_y_continuous(labels = comma)

```

2. Create a scatter plot of Sales vs. Profit with a trend line.
```{r}
ggplot(super_store, aes(x = Sales, y = Profit)) +
  geom_point(color = "blue", size = 2) +
  geom_smooth(method = "lm", color = "red", se = TRUE) +
  labs(title = "Scatter Plot of Sales vs. Profit with Trend Line",
       x = "Sales",
       y = "Profit") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
  
```

3. Create a histogram of Profit_Margin.
```{r}
#Histogram of Profit Margin

ggplot(data = super_store, aes(x = Profit_Margin)) +
  geom_histogram(binwidth = 0.05, fill = "steelblue", color = "black") +
  labs(title = "Distribution of Profit Margin",
       x = "Profit Margin",
       y = "Frequency") +
  theme_minimal()+
  theme(plot.title = element_text(hjust = 0.5))
        
```



